---
title: "Baseline and MR generation (v02)"
author: "Gabriel"
date: "2026-05-19"
format:
  html:
    theme: darkly
    toc: true
    code-fold: false
    embed-resources: true
---

Simulates the 0% MR baseline with `remulateTie`, then injects MR at 0/2/8/16/24%
via additive random tagging. Writes CSVs to `code/data/synthetic/mr_levels_v02/`.

## Setup

```{r}
#| label: setup
#| message: false

library(remulate)

project_root <- normalizePath(file.path("..", ".."), mustWork = TRUE)
output_dir   <- file.path(project_root, "code", "data", "synthetic")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
```

## Baseline (0% MR)

```{r}
#| label: baseline-generation

set.seed(20260423)

n_actors <- 20
n_events <- 2000

actor_attrs <- data.frame(
  id = 1:n_actors,
  time = rep(0, n_actors),
  group = sample(c(0, 1), n_actors, replace = TRUE),
  active = rnorm(n_actors),
  senior = sample(c(0, 1), n_actors, replace = TRUE, prob = c(0.7, 0.3))
)

effects_baseline <- ~ baseline(-5) +
  inertia(0.02) +
  reciprocity(0.10) +
  otp(0.03) +
  difference(0.15, variable = "active", attr_actors = actor_attrs) +
  same(0.4, variable = "group", attr_actors = actor_attrs)

reh_0mr <- remulateTie(
  effects = effects_baseline,
  actors  = 1:n_actors,
  endTime = Inf,
  events  = n_events,
  initial = 0
)

reh_0mr$event_id <- seq_len(nrow(reh_0mr))

stopifnot(sum(duplicated(reh_0mr$time)) == 0)
```

```{r}
#| label: save-baseline

write.csv(reh_0mr,
          file.path(output_dir, "reh_baseline_0mr_20actors_2000events.csv"),
          row.names = FALSE)
write.csv(actor_attrs,
          file.path(output_dir, "reh_baseline_0mr_actor_attributes.csv"),
          row.names = FALSE)
```

## MR helper (additive random tagging)

We turn the one-row-per-event baseline into a "long" table with one row per
(sender, receiver) pair. A random fraction `p_mr_event` of events is tagged as
multi-receiver, and each tagged event gets one extra receiver (chosen so it is
not the sender or the original receiver).

```{r}
#| label: mr-v02-helper

make_mr <- function(reh_df, actor_ids, p_mr_event, seed) {

  set.seed(seed)
  n_events <- nrow(reh_df)
  n_to_tag <- round(p_mr_event * n_events)

  # randomly pick which events become multi-receiver
  tagged_ids <- sample(reh_df$event_id, n_to_tag)

  # build the long table one event at a time
  mr_long <- data.frame()
  for (i in 1:n_events) {
    row_i <- reh_df[i, ]
    is_tagged <- row_i$event_id %in% tagged_ids
    this_size <- if (is_tagged) 2 else 1

    # every event keeps its original sender -> receiver row
    base_row <- data.frame(
      event_id = row_i$event_id,
      sender   = row_i$sender,
      receiver = row_i$receiver,
      time     = row_i$time,
      mr_size  = this_size
    )
    mr_long <- rbind(mr_long, base_row)

    # tagged events also get one extra receiver
    if (is_tagged) {
      candidates <- setdiff(actor_ids, c(row_i$sender, row_i$receiver))
      extra_receiver <- sample(candidates, 1)
      extra_row <- data.frame(
        event_id = row_i$event_id,
        sender   = row_i$sender,
        receiver = extra_receiver,
        time     = row_i$time,
        mr_size  = this_size
      )
      mr_long <- rbind(mr_long, extra_row)
    }
  }

  mr_long <- mr_long[order(mr_long$event_id), ]

  # count how many receivers each event ended up with
  mr_event_table <- aggregate(receiver ~ event_id + sender + time + mr_size,
                              data = mr_long, FUN = length)
  names(mr_event_table)[names(mr_event_table) == "receiver"] <- "n_receivers_observed"
  mr_event_table <- mr_event_table[order(mr_event_table$event_id), ]

  list(
    mr_long          = mr_long,
    mr_event_table   = mr_event_table,
    achieved_p_event = n_to_tag / n_events,
    n_mr_events      = n_to_tag,
    n_long_rows      = nrow(mr_long),
    seed             = seed
  )
}
```

## Build MR levels

```{r}
#| label: build-mr-levels-v02

mr_levels  <- c(0, 0.02, 0.08, 0.16, 0.24)
mr_dir_v02 <- file.path(output_dir, "mr_levels_v02")
dir.create(mr_dir_v02, recursive = TRUE, showWarnings = FALSE)

mr_master_seed <- 20260519

# one entry per MR level
mr_results <- list()

for (p in mr_levels) {
  seed_k <- mr_master_seed + as.integer(p * 1000)

  res <- make_mr(
    reh_df     = reh_0mr,
    actor_ids  = 1:n_actors,
    p_mr_event = p,
    seed       = seed_k
  )

  # e.g. p = 0.08 -> "mr08"
  level_tag <- sprintf("mr%02d", as.integer(round(p * 100)))
  mr_results[[level_tag]] <- res

  write.csv(
    res$mr_long,
    file.path(mr_dir_v02, paste0("reh_", level_tag, "_from_baseline_long.csv")),
    row.names = FALSE
  )
  write.csv(
    res$mr_event_table,
    file.path(mr_dir_v02, paste0("reh_", level_tag, "_from_baseline_event_table.csv")),
    row.names = FALSE
  )
}
```

## Checks

```{r}
#| label: verify-v02

# build one summary row per MR level
check_df <- data.frame()
for (p in mr_levels) {
  level_tag <- sprintf("mr%02d", as.integer(round(p * 100)))
  res <- mr_results[[level_tag]]
  check_df <- rbind(check_df, data.frame(
    level            = paste0(round(p * 100), "%"),
    target_p_event   = p,
    mr_event_count   = res$n_mr_events,
    achieved_p_event = round(res$achieved_p_event, 4),
    long_rows        = res$n_long_rows,
    seed             = res$seed
  ))
}

check_df
```

```{r}
#| label: receiver-uniqueness-check

# sanity check: within each multi-receiver event the receivers should all be
# different from each other and different from the sender
for (p in mr_levels) {
  level_tag <- sprintf("mr%02d", as.integer(round(p * 100)))
  mr_long <- mr_results[[level_tag]]$mr_long
  mr_only <- mr_long[mr_long$mr_size > 1, ]
  if (nrow(mr_only) == 0) next

  for (id in unique(mr_only$event_id)) {
    grp <- mr_only[mr_only$event_id == id, ]
    duplicate_receiver <- any(duplicated(grp$receiver))
    receiver_is_sender <- any(grp$receiver == grp$sender[1])
    if (duplicate_receiver || receiver_is_sender) {
      stop(paste0("Receiver uniqueness violated at MR = ", p * 100,
                  "% for event ", id))
    }
  }
}
```
