---
title: "REM Parameter Recovery on 0% MR Baseline (v01)"
author: "Gabriel"
date: "2026-04-30"
format:
  html:
    theme: darkly
    toc: true
    code-fold: false
    embed-resources: true
---

## Purpose

This notebook validates the REM estimation pipeline by fitting a tie-oriented
REM to the clean **0% MR baseline** and checking whether the estimated
parameters recover the known generating values.

This is the gating step before any multi-receiver experiments (Decision #8 in
`THESIS_TRACKER.md`): if the pipeline cannot recover the truth from clean
dyadic data, downstream comparisons are not trustworthy.

**True generating parameters** (from `baseline_and_mr_generation_v02.qmd`):

| Statistic          | Type       | True value |
|--------------------|------------|-----------|
| baseline           | —          | -5.00     |
| inertia            | endogenous | 0.02      |
| reciprocity        | endogenous | 0.10      |
| otp                | endogenous | 0.03      |
| difference (active)| exogenous  | 0.15      |
| same (group)       | exogenous  | 0.40      |

## 1) Setup

```{r}
#| label: setup
#| message: false

library(remify)
library(remstats)
library(remstimate)

# project_root points to the repo root (three folders up from this notebook)
project_root <- normalizePath(file.path("..", "..", ".."), mustWork = TRUE)
```

## 2) Load baseline data

```{r}
#| label: load-data

data_dir <- file.path(project_root, "code", "data", "synthetic")

reh_raw <- read.csv(file.path(data_dir, "reh_baseline_0mr_20actors_2000events.csv"))
actor_attrs <- read.csv(file.path(data_dir, "reh_baseline_0mr_actor_attributes.csv"))
```

Loaded `r nrow(reh_raw)` events on `r nrow(actor_attrs)` actors. Columns:
`r paste(names(reh_raw), collapse = ", ")`.

```{r}
#| label: load-data-preview
head(reh_raw)
head(actor_attrs)
```

## 3) Process with remify

`remify::remify()` expects the first three columns of the edgelist to be
`[time, actor1, actor2]` in that order. Our CSV already has
`time, sender, receiver` as the first three — column names do not matter,
only position.

```{r}
#| label: remify

edgelist <- reh_raw[, c("time", "sender", "receiver")]

reh <- remify::remify(
  edgelist = edgelist,
  directed = TRUE,
  ordinal  = FALSE,
  model    = "tie",
  actors   = 1:20
)
```

## 4) Compute statistics with remstats

The formula mirrors the generating specification but **without coefficient
values** — those are what we estimate. Actor attributes require a `name`
column that matches the actor identifiers used in the REH.

```{r}
#| label: remstats

actor_attrs_rs <- data.frame(
  name   = as.character(actor_attrs$id),
  time   = actor_attrs$time,
  group  = actor_attrs$group,
  active = actor_attrs$active
)

tie_effects <- ~ 1 +
  inertia() +
  reciprocity() +
  otp() +
  difference(variable = "active") +
  same(variable = "group")

stats <- remstats::remstats(
  reh         = reh,
  tie_effects = tie_effects,
  attr_actors = actor_attrs_rs
)
```

## 5) Estimate with remstimate (MLE)

```{r}
#| label: remstimate

fit <- remstimate::remstimate(
  reh    = reh,
  stats  = stats,
  method = "MLE"
)
```

## 6) Parameter recovery comparison

```{r}
#| label: recovery

true_params <- c(
  baseline = -5.00,
  inertia = 0.02,
  reciprocity = 0.10,
  otp = 0.03,
  difference_active = 0.15,
  same_group = 0.40
)

# estimated coefficients and their standard errors
estimated <- coef(fit)
se_vec <- sqrt(diag(vcov(fit)))

recovery <- data.frame(
  statistic = names(true_params),
  true_value = true_params,
  estimated = estimated,
  se = se_vec,
  difference = estimated - true_params,
  row.names = NULL
)

# a parameter is "recovered" if the estimate is within 2 SE of the truth
recovery$within_2se <- abs(recovery$difference) < 2 * recovery$se

recovery
```

## 7) Summary

```{r}
#| label: summary

data.frame(
  max_abs_deviation  = max(abs(recovery$difference)),
  mean_abs_deviation = mean(abs(recovery$difference)),
  n_within_2se       = sum(recovery$within_2se, na.rm = TRUE),
  n_statistics       = nrow(recovery),
  all_within_2se     = all(recovery$within_2se, na.rm = TRUE)
)
```
Max absolute deviation: **`r round(max(abs(recovery$difference)), 4)`**;
mean absolute deviation: **`r round(mean(abs(recovery$difference)), 4)`**.
Parameters within 2 SE of truth: **`r sum(recovery$within_2se, na.rm = TRUE)` /
`r nrow(recovery)`**.