---
title: "RHEM / eventnet prep (v01)"
author: "Gabriel"
date: "2026-04-30"
format:
  html:
    theme: darkly
    toc: true
    code-fold: false
    embed-resources: true
---

Export MR datasets for eventnet, then fit RHEM with `coxph` on the design matrices.

Inputs: `mr_levels_v02/` from `baseline_and_mr_generation_v02.qmd`.  
Outputs: CSVs in `eventnet_exports/`; coefficients in `eventnet_gui_statistics/rhem_coefficient_summary.csv`.

After exporting, run eventnet on each CSV (see `eventnet_gui_statistics/eventnet.configuration.txt`) and save design matrices to `eventnet_gui_statistics/`.

## Setup

```{r}
#| label: setup
#| message: false

library(survival)

project_root <- normalizePath(file.path("..", "..", ".."), mustWork = TRUE)
mr_dir       <- file.path(project_root, "code", "data", "synthetic", "mr_levels_v02")
export_dir   <- file.path(project_root, "code", "data", "synthetic", "eventnet_exports")
stats_dir    <- file.path(project_root, "code", "data", "synthetic", "eventnet_gui_statistics")
dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)
```

## Load MR data

```{r}
#| label: load-data

mr_levels <- c(0.02, 0.08, 0.16, 0.24)

mr_long_list <- list()
for (p in mr_levels) {
  tag <- sprintf("mr%02d", as.integer(round(p * 100)))
  mr_long_list[[tag]] <- read.csv(
    file.path(mr_dir, paste0("reh_", tag, "_from_baseline_long.csv"))
  )
}

actor_attrs <- read.csv(
  file.path(project_root, "code", "data", "synthetic",
            "reh_baseline_0mr_actor_attributes.csv")
)
```

## Build eventnet CSVs

```{r}
#| label: build-eventnet-csvs

n_actors <- nrow(actor_attrs)

# eventnet wants the actor attributes to be "set" at a time before the first
# real event, so we use a dummy time just below the earliest event time
t_dummy <- min(mr_long_list[[1]]$time) - 1

build_eventnet_csv <- function(long_df, actor_df, t_dummy) {
  # one "add.actor" row per actor at the dummy time
  add_actors <- data.frame(
    message.id = 0, sender.id = actor_df$id, receiver.id = actor_df$id,
    time = t_dummy, type = "add.actor", weight = 1
  )

  # group membership: actors with group == 1
  group_ids <- actor_df$id[actor_df$group == 1]
  attr_group <- data.frame(
    message.id = 0, sender.id = group_ids, receiver.id = group_ids,
    time = t_dummy, type = "is.group", weight = 1
  )

  # the continuous "active" attribute is stored in the weight column
  attr_active <- data.frame(
    message.id = 0, sender.id = actor_df$id, receiver.id = actor_df$id,
    time = t_dummy, type = "active", weight = actor_df$active
  )

  # the real email events
  events <- data.frame(
    message.id = long_df$event_id, sender.id = long_df$sender,
    receiver.id = long_df$receiver, time = long_df$time,
    type = "email", weight = 1
  )

  out <- rbind(add_actors, attr_group, attr_active, events)
  out[order(out$time, out$message.id), ]
}

for (tag in names(mr_long_list)) {
  out <- build_eventnet_csv(mr_long_list[[tag]], actor_attrs, t_dummy)
  write.csv(out, file.path(export_dir, paste0(tag, "_eventnet.csv")),
            row.names = FALSE)
}
```

## Verify exports

```{r}
#| label: verify

verify_df <- data.frame()
for (f in list.files(export_dir, pattern = "\\.csv$")) {
  df <- read.csv(file.path(export_dir, f))
  verify_df <- rbind(verify_df, data.frame(
    file = f,
    event_rows = sum(df$type == "email"),
    dummy_rows = sum(df$type != "email")
  ))
}

verify_df
```

## RHEM estimation (coxph)

Expects eventnet outputs named `{tag}_eventnet_{TAG}_COND_SENDER.csv` in `stats_dir`.

```{r}
#| label: rhem-estimation

rhem_results <- list()
mr_tags <- sprintf("mr%02d", as.integer(round(mr_levels * 100)))

for (tag in mr_tags) {
  obs_name <- paste0(toupper(tag), "_COND_SENDER")
  out_file <- file.path(stats_dir, paste0(tag, "_eventnet_", obs_name, ".csv"))
  if (!file.exists(out_file)) next

  events <- read.csv(out_file)

  # keep the real events plus the non-observed risk-set rows eventnet added
  events <- events[events$TYPE == "email" | events$IS_OBSERVED == 0, ]

  # the Enron tutorial applies sqrt() to the endogenous statistics before fitting
  for (v in c("s.r.sub.rep.1", "reciprocation", "otp")) {
    if (v %in% names(events)) events[[v]] <- sqrt(events[[v]])
  }

  rhem_results[[tag]] <- coxph(
    Surv(time = rep(1, nrow(events)), event = events$IS_OBSERVED) ~
      s.r.sub.rep.1 + reciprocation + otp + absdiff.active + catdiff.group +
      strata(EVENT_INTERVAL),
    data = events
  )
}

length(rhem_results)
```

## Coefficient summary

```{r}
#| label: extract-coefs

# translate eventnet's statistic names into our generating-effect names
effect_map <- c(
  "s.r.sub.rep.1"  = "inertia",
  "reciprocation"  = "reciprocity",
  "otp"            = "otp",
  "absdiff.active" = "difference_active",
  "catdiff.group"  = "same_group"
)

rhem_coefs <- data.frame()
for (tag in names(rhem_results)) {
  fit <- rhem_results[[tag]]
  stat_names <- names(coef(fit))
  rhem_coefs <- rbind(rhem_coefs, data.frame(
    mr_level  = tag,
    statistic = stat_names,
    estimate  = as.numeric(coef(fit)),
    se        = sqrt(diag(vcov(fit))),
    effect    = effect_map[stat_names]
  ))
}

if (nrow(rhem_coefs) > 0) {
  write.csv(rhem_coefs, file.path(stats_dir, "rhem_coefficient_summary.csv"),
            row.names = FALSE)
}

rhem_coefs
```
