---
title: "Noise-and-Reshuffle REM Estimation Across MR Levels (v02)"
author: "Gabriel"
date: "2026-06-03"
format:
  html:
    theme: darkly
    toc: true
    code-fold: false
    embed-resources: true
---

## 1) Setup

```{r}
#| label: setup
#| message: false

library(remify)
library(remstats)
library(remstimate)

project_root <- normalizePath(file.path("..", "..", ".."), mustWork = TRUE)
data_dir   <- file.path(project_root, "code", "data", "synthetic")
mr_dir     <- file.path(data_dir, "mr_levels_v02")
output_dir <- file.path(data_dir, "rem_noise_estimates")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

output_csv <- file.path(output_dir, "rem_noise_estimates_v02.csv")
log_csv    <- file.path(output_dir, "rem_noise_run_log_v02.csv")
```

## 2) Configuration

```{r}
#| label: config

R_REPLICATES  <- 5    # repeat each MR level a few times as a robustness check
MR_LEVELS_PCT <- c(0, 2, 8, 16, 24)
EPS_FACTOR    <- 0.10  # noise size = 0.1 * smallest gap between event times
MASTER_SEED   <- 20260521
N_ACTORS      <- 20

# the values used by the simulator (what we hope to recover)
TRUE_PARAMS <- c(
  baseline = -5.00,
  inertia = 0.02,
  reciprocity = 0.10,
  otp = 0.03,
  difference_active = 0.15,
  same_group = 0.40
)
```

## 3) Load baseline and MR-level data

Every MR level is derived from the same 0% baseline
(see `baseline_and_mr_generation_v02.qmd`). For the 0% level we build a matching
long table here so the fitting code can treat all levels the same way.

```{r}
#| label: load-data

actor_attrs <- read.csv(file.path(data_dir, "reh_baseline_0mr_actor_attributes.csv"))

reh_0mr <- read.csv(file.path(data_dir, "reh_baseline_0mr_20actors_2000events.csv"))
reh_0mr_long <- data.frame(
  event_id = reh_0mr$event_id,
  sender   = reh_0mr$sender,
  receiver = reh_0mr$receiver,
  time     = reh_0mr$time,
  mr_size  = 1L
)

mr_long_data <- list("0" = reh_0mr_long)
for (pct in MR_LEVELS_PCT[MR_LEVELS_PCT > 0]) {
  tag <- sprintf("mr%02d", pct)
  mr_long_data[[as.character(pct)]] <- read.csv(
    file.path(mr_dir, paste0("reh_", tag, "_from_baseline_long.csv"))
  )
}

load_summary <- data.frame()
for (pct in MR_LEVELS_PCT) {
  df <- mr_long_data[[as.character(pct)]]
  load_summary <- rbind(load_summary, data.frame(
    mr_level_pct         = pct,
    rows                 = nrow(df),
    unique_events        = length(unique(df$event_id)),
    duplicate_timestamps = sum(duplicated(df$time))
  ))
}

load_summary
```

## 4) Helper functions

### 4a) Noise injection

A tie-oriented REM needs a strict event order, but multi-receiver events all
share the same timestamp. `inject_noise()` deals with this by adding a tiny
random amount of time to every row that shares its timestamp with another row
(all MR rows, plus the occasional tied baseline event).

The noise is drawn from `Uniform(0, eps_max)`, where `eps_max` is `eps_factor`
times the smallest gap between distinct event times. Keeping the noise smaller
than the smallest real gap means a perturbed event can never jump past a
neighbouring event, so the order of the distinct timestamps is preserved.

```{r}
#| label: inject-noise

inject_noise <- function(long_df, eps_factor, seed) {

  # which rows share their timestamp with at least one other row?
  shared_time <- duplicated(long_df$time) | duplicated(long_df$time, fromLast = TRUE)
  needs_noise <- (long_df$mr_size > 1) | shared_time

  # biggest noise we allow: a fraction of the smallest gap between event times
  unique_times <- sort(unique(long_df$time))
  if (length(unique_times) < 2) {
    eps_max <- 0
  } else {
    eps_max <- min(diff(unique_times)) * eps_factor
  }

  # add noise only to the rows that need it
  set.seed(seed)
  noise <- numeric(nrow(long_df))
  if (eps_max > 0) {
    noise[needs_noise] <- runif(sum(needs_noise), min = 0, max = eps_max)
  }

  long_df$noise_added <- noise
  long_df$time_noisy  <- long_df$time + noise
  long_df <- long_df[order(long_df$time_noisy, long_df$event_id), ]

  # the noisy times should now all be distinct
  stopifnot(!anyDuplicated(long_df$time_noisy))

  list(
    df        = long_df,
    eps_max   = eps_max,
    n_mr_rows = sum(long_df$mr_size > 1),
    n_noised  = sum(needs_noise)
  )
}
```

### 4b) REM fitter

Same effects formula and MLE estimator as `rem_parameter_recovery_v01.qmd`.
It takes the noisy long-format edgelist and returns the coefficients, their
standard errors, and how long the fit took.

```{r}
#| label: rem-fitter

REM_EFFECTS <- ~ 1 +
  inertia() +
  reciprocity() +
  otp() +
  difference(variable = "active") +
  same(variable = "group")

actor_attrs_rs <- data.frame(
  name   = as.character(actor_attrs$id),
  time   = actor_attrs$time,
  group  = actor_attrs$group,
  active = actor_attrs$active
)

fit_rem_on_long <- function(long_df, actors, effects, attrs_rs) {

  edgelist <- data.frame(
    time     = long_df$time_noisy,
    actor1   = long_df$sender,
    actor2   = long_df$receiver
  )

  t_start <- Sys.time()

  reh <- remify::remify(
    edgelist = edgelist,
    directed = TRUE,
    ordinal  = FALSE,
    model    = "tie",
    actors   = actors
  )

  stats <- remstats::remstats(
    reh         = reh,
    tie_effects = effects,
    attr_actors = attrs_rs
  )

  fit <- remstimate::remstimate(
    reh    = reh,
    stats  = stats,
    method = "MLE"
  )

  t_end <- Sys.time()

  list(
    estimate    = coef(fit),
    se          = sqrt(diag(vcov(fit))),
    elapsed_sec = as.numeric(difftime(t_end, t_start, units = "secs"))
  )
}
```

### 4c) One-replicate driver

```{r}
#| label: one-replicate

run_one_replicate <- function(long_df, mr_pct, replicate_id,
                              eps_factor, master_seed,
                              actors, effects, attrs_rs) {

  seed <- as.integer(master_seed + mr_pct * 1000L + replicate_id)
  injected <- inject_noise(long_df, eps_factor = eps_factor, seed = seed)
  fit_out  <- fit_rem_on_long(
    long_df  = injected$df,
    actors   = actors,
    effects  = effects,
    attrs_rs = attrs_rs
  )

  data.frame(
    mr_level_pct = mr_pct,
    replicate    = replicate_id,
    seed         = seed,
    eps_factor   = eps_factor,
    eps_max      = injected$eps_max,
    n_mr_rows    = injected$n_mr_rows,
    n_noised     = injected$n_noised,
    n_rows       = nrow(injected$df),
    statistic    = names(fit_out$estimate),
    estimate     = as.numeric(fit_out$estimate),
    se           = as.numeric(fit_out$se),
    elapsed_sec  = fit_out$elapsed_sec,
    row.names    = NULL,
    stringsAsFactors = FALSE
  )
}
```

## 5) Main loop — every MR level × R replicates

This fits one REM per (MR level, replicate) and writes the results to a
long-format CSV.

```{r}
#| label: main-loop

rem_results <- data.frame()
run_log     <- data.frame()

for (pct in MR_LEVELS_PCT) {
  long_df <- mr_long_data[[as.character(pct)]]

  for (r in 1:R_REPLICATES) {
    one <- run_one_replicate(
      long_df      = long_df,
      mr_pct       = pct,
      replicate_id = r,
      eps_factor   = EPS_FACTOR,
      master_seed  = MASTER_SEED,
      actors       = 1:N_ACTORS,
      effects      = REM_EFFECTS,
      attrs_rs     = actor_attrs_rs
    )
    rem_results <- rbind(rem_results, one)

    # one log row per replicate (all rows of `one` share these values)
    run_log <- rbind(run_log, data.frame(
      mr_level_pct = pct,
      replicate    = r,
      seed         = one$seed[1],
      eps_max      = one$eps_max[1],
      elapsed_sec  = one$elapsed_sec[1]
    ))
  }
}

write.csv(rem_results, output_csv, row.names = FALSE)
write.csv(run_log,     log_csv,    row.names = FALSE)

loop_msg <- paste0(
  "Ran ", R_REPLICATES * length(MR_LEVELS_PCT),
  " REM fits and saved results to ", basename(output_csv), "."
)
```

`r loop_msg`

## 6) Summary across MR levels

For each MR level we average the estimates over the replicates and compare the
mean to the true value. The 0% level is the noise-free baseline, so its
estimates should match `rem_parameter_recovery_v01.qmd`.

```{r}
#| label: cell-summary

# average each statistic across replicates, separately for every MR level
combos <- unique(rem_results[, c("mr_level_pct", "statistic")])

agg <- data.frame()
for (i in 1:nrow(combos)) {
  pct  <- combos$mr_level_pct[i]
  stat <- combos$statistic[i]
  vals <- rem_results$estimate[rem_results$mr_level_pct == pct &
                                 rem_results$statistic == stat]
  agg <- rbind(agg, data.frame(
    mr_level_pct  = pct,
    statistic     = stat,
    mean_estimate = mean(vals),
    sd_estimate   = sd(vals),
    n_replicates  = length(vals)
  ))
}

# attach the true value and the bias (mean estimate - truth)
agg$true_value <- TRUE_PARAMS[agg$statistic]
agg$bias <- agg$mean_estimate - agg$true_value

agg <- agg[order(agg$mr_level_pct, agg$statistic), ]
write.csv(agg, file.path(output_dir, "rem_noise_cell_summary_v02.csv"),
          row.names = FALSE)
agg
```
